\documentclass[12pt]{article}

\usepackage{fontspec}
\usepackage{geometry}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage{hyperref}

\geometry{top=1in, bottom=1in, left=1in, right=1in, marginparsep=4pt, marginparwidth=1in}

\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancyplain}
\fancyhf{}
\cfoot{\thepage\ of \pageref{LastPage}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{12pt}

% \setromanfont [Ligatures={Common}, Numbers={OldStyle}, Variant=01,
%  BoldFont={LinLibertine_RB.otf},
%  ItalicFont={LinLibertine_RI.otf},
%  BoldItalicFont={LinLibertine_RBI.otf}
%  ]{LinLibertine_R.otf}

\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

\usepackage{xunicode}
\defaultfontfeatures{Mapping=tex-text}

\setromanfont{YaleNew}

\begin{document}

\begin{center}
{\bf Data Collection Assignment -- Lab 15} \\
\end{center}

\bigskip

\textbf{Overview:} \vspace{6pt}

Remember at the start of the semester how I said we would have
class outside once? If you are reading this, today is the day! We
are going to run around outside collecting images that we can
use in class once we start analysing image data.

Completing this assignment counts as `Lab 15'. There will not be
a separate task to finish.

\textbf{General Instructions:} \vspace{6pt}

Your goal is to take a set of at least $60$ photos. There should be
a minimum of $15$ images in each of these categories:
\begin{itemize}
\item images taken outdoors without people in the foreground
\item images taken outdoors with someone from our class in the foreground
\item images taken indoors without people in the foreground
\item images taken indoors with someone from our class in the foreground
\end{itemize}
Please make sure to take a variety of shots from across campus.
Don't take more than one from any particular place. For the indoor
photos, this means only one photo of each type (with and without someone
else) in a particular building.

You will need to work in groups for this to work because you
need images with other students. I suggest having groups of no more
than 5 people. However, each student needs to have their own set
of $60$ images. My goal is end up with a dataset of around 1400 images.

\textit{If you do not have a phone with a camera, please let me know
before you leave so we can figure out the best plan of action. I may
have a spare camera or can let you partner with someone else.}

\textbf{Uploading:} \vspace{6pt}

I have invited you all to join a Box folder called \textbf{stat395\_images}.
You should see a directory with your GitHub name attached to it. Inside are
four more directories corresponding to the four categories of images. Please
upload your images to the relevant directory before our next class.

Note that this directory is open for the entire class to see. We will be
working with these images as a class in a future lab. If you feel uncomfortable
sharing your images publicly, please let me know.

\end{document}






