---
title: "Class 13: Congratulations, You've Won $1M!"
author: "Taylor Arnold"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/2017-10-10-class13/")
```

```{r, message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(methods)
library(stringi)
library(smodels)
library(tokenizers)
```

## Spam Detection

Today, we start our analysis of textual data (to me, this is where things
start to get particularly interesting). As a first task let's look at a
dataset of spam records:

```{r, message = FALSE}
spam <- read_csv("https://statsmaths.github.io/ml_data/spam.csv")
```

Notice that our dataset now has only four columns. There is no easy way of
putting this in any of our predictive models.

Its useful to see some examples of the data, particularly to see if we can
detect any manual patterns for classifying the data. What do you think the
classification of this message is?

```{r}
stri_wrap(spam$text[500], width = 60)
```

Or this one:

```{r}
stri_wrap(spam$text[506], width = 60)
```

Or this:

```{r}
stri_wrap(spam$text[903], width = 60)
```

Or this:

```{r}
stri_wrap(spam$text[125], width = 60)
```

As you probably guessed, the first and last are spam and the other two
are not:

```{r}
spam$class[c(500, 506, 903, 125)]
```

## Constructing Features by Hand

In order to build predictive models from this dataset, we first need to
create new (ideally numeric) variables that capture relevant aspects of
each message. This process is known as *featurization*, and will be the
key task that we will focus on in our text analysis.

What things might be useful to extract from the text? In a moment, we'll
see some more systematic ways of doing this, but for spam we can do some
of this manually. Perhaps we could detect the length of each text and use
that as a feature? To do that, I'll use the **stringi** function `stri_length`:

```{r}
spam$length <- stri_length(spam$text)
```

It may also be useful to count how many times the pound symbol (this data
comes from the UK) and exclamation symbol is used. For this, we use the
`stri_count` function:

```{r}
spam$count_pound <- stri_count(spam$text, fixed = "£")
spam$count_exclm <- stri_count(spam$text, fixed = "!")
```

We can tweak the count function to match patterns rather than a fixed
string. Let's count the number of capital letters and the number of
numbers:

```{r}
spam$count_caps <- stri_count(spam$text, regex = "[A-Z]")
spam$count_nums <- stri_count(spam$text, regex = "[0-9]")
```

The `regex` option stands for *regular expression*. We are not going to
make much use of them in this course, but they are incredibly powerful
when cleaning and manipulating text-based datasets.

With these new variables in hand, we can now apply the learning algorithms
we learned earlier this semester. I'll use a GAM model, allowing for a
non-linear relationship with length.

```{r}
library(gam)
model <- gam(class ~ s(length) + count_pound + count_exclm + count_nums +
                     count_caps,
             data = spam,
             subset = (train_id == "train"))
plot(model)
```

As we may have expected, all of the linear terms are positively related
with spam. Spam tends to have about 200 characters, with the probability
decreasing for shorter and longer messages.

```{r}
spam$class_pred <- as.numeric(predict(model, newdata = spam, type = "response") > 0.5)
tapply(spam$class == spam$class_pred, spam$train_id, mean)
```

So, we are already doing a very good job with about a 92% classification
rate. Here, a confusion matrix is fairly useful:

```{r}
table(y = spam$class, ypred = spam$class_pred)
```

It shows that only classify three non-spam messages as spam but fail
to detect 76 spam messages. This inbalance, where one type of error is
more common, comes up a lot in text and image analysis. Knowing which
error type is more frequent is helpful in figuring what features are
missing.

## Negative examples

Let's look 50 of those elements in the training sampling that were not
detected as spam. Can you think of things that would have helped
or why the model was not correctly classifying the output?

```{r}
these_rows <- which(spam$train_id == "train" &
                    spam$class != spam$class_pred &
                    spam$class == 1)
these_rows <- sample(these_rows, 50)
cat(stri_wrap(spam$text[these_rows], exdent = 5, width = 60), sep = "\n")
```

In some cases different model might help (there are some short messages
that use all of the other "spam" features that would probably be caught by
a local model).

## Product Classification

The dataset for your next lab has a similar task, but involves multiclass
classification. The task is to look at amazon reviews, which we can read
in with the following:

```{r, message = FALSE}
amazon <- read_csv("https://statsmaths.github.io/ml_data/amazon_product_class.csv")
```

And to classify them as being in:

1. Books
2. Food
3. Movies and TV

Here are some examples:

```{r}
these_rows <- which(amazon$train_id == "train" &
                    !duplicated(amazon$category))
cat(stri_wrap(amazon$text[these_rows], exdent = 5, width = 60), sep = "\n")
```

**What might make these easier to classifier and what might make them
hard to classifier?**

### Detecting words

To do a good job of predicting categories in the Amazon dataset, we need
to create features that capture how frequently certain words are used.
For example, it might make sense to figure out how often the word "read"
was used in each review. The thought is that it is common in the
book reviews (and likely entirely absent from food reviews).

```{r}
amazon$count_read <- stri_count(amazon$text, fixed = "read")
```

With three categories and only a small number of features, the `multinom`
function is a good choice for a model to fit this feature to:

```{r}
library(nnet)
model <- multinom(category ~ count_read, data = amazon)
```

Remembering that the base class is the book category, the slopes
of the read feature (which are negative here) seem as expected:

```{r}
coef(model)
```

We see that, on its own, this one feature lets us do a reasonably good
job of prediction (remember to compare it to the 33% rate of random
guessing):

```{r}
amazon$category_pred <- predict(model, newdata = amazon)
tapply(amazon$category_pred == amazon$category, amazon$train_id, mean)
```

With this feature we can start to see the benefit of the confusion
matrix:

```{r}
table(y = amazon$category, y_pred = amazon$category_pred)
```

Notice that our simple model never predicts that something is a Food
item and only predicts that it is a Book or a Movie. It's main predictive
power coems from seperating some Books from the other categories, as we
would have expected.

For the next lab, your task is to create more features by finding other
words that help to seperate the classes. This might be due to just thinking
about the problem, as we did here, or looking at negative examples to find
patterns. Some other things you might try:

- use the function `stri_trans_tolower` to convert the text to lower
case before counting tokens
- does it make more sense to normalize each county by the length of
the review?
- is there any predictive power to features like length or other stylistic
features
- notice that searching for "read" also detects words that contain "read";
this might be useful such as with "reading" or less useful which words
such as "ready"

There is a lot of room to be creative this time, but don't feel the need
to go too crazy. As we did with structured data in the first few weeks,
we will see more automatic/systematic ways of doing this analysis in the
next class.








