---
title: "Class 12: Taxi!"
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/2017-10-05-class12/")
```

```{r, message = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(methods)
```

## Taxi Data

Today I am going to look at a dataset of taxi rides in NYC. For each
ride we want to predict how long the journey took:

```{r, message = FALSE}
taxi <- read_csv("https://statsmaths.github.io/ml_data/nyc_taxi.csv")
```

Notice that we have only a limited amount of information about each
ride:

- time: day, hour, minute, and weekday
- location: pickup and dropoff latitude / longitude / neighborhood
- derived variables: trip distance

As way of review, we'll go over all of the primary estimators we have 
learned and show how they might be applied to this prediction problem.

One key thing here is to learn how we would approach a problem like
this. Models should usually not be applied in the order we learned
them.

## Model Selection and Baseline

My first step is to figure out which variables might be of interest and to
establish a quick baseline for how well I can generally expect to estimate
the output. Here we have only 11 variables, and only two are directly 
categorical. This is a borderline case between using **xgboost** and 
**glmnet**, so let's use both

To use the either for model selection, I will usually put all of the
variables into the model. I leave out only very grainular categorical
variables (such as the player name in the NBA shots data). 

```{r}
X <- model.matrix(~ . - 1, data = taxi[,4:14])
y <- taxi$duration

X_train <- X[taxi$train_id == "train",]
y_train <- y[taxi$train_id == "train"]
X_valid <- X[taxi$train_id == "valid",]
y_valid <- y[taxi$train_id == "valid"]
```

### Elastic Net

Now, we can apply the elastic net. Setting alpha equal to 0.9 is a good
starting spot.

```{r, message = FALSE}
library(glmnet)
model <- cv.glmnet(X_train, y_train, alpha = 0.9)
taxi$duration_pred_enet <- predict(model, newx = X)
beta <- coef(model, s = c(model$lambda.min, model$lambda.1se, model$lambda[c(3, 10)]))
beta[beta[,2] != 0,]
```

Above, I printed out only those rows where the 1se model has non-zero coefficents.
This limits the output only to those most important variables. For interest, I also
added several larger lambdas. Here is the predictiveness of the model:

```{r}
sqrt(tapply((taxi$duration_pred_enet - taxi$duration)^2, taxi$train_id, mean))
```

What do we learn from the elastic net? Here are a summary of things learned:

- Trip distance is the most important variable (unsurprising, perhaps)
- The dropoff location effects the output significantly
- The hour and day of the week also has an effect
- We should be able to get an RMSE of at least around 370 seconds on the validation set

### Boosted Trees

We can repeat this analysis of the best variables with boosted trees. These
may find different variable importance scores because it allows for learned
non-linearities and interactions.

```{r}
library(xgboost)
data_train <- xgb.DMatrix(data = X_train, label = y_train)
data_valid <- xgb.DMatrix(data = X_valid, label = y_valid)
watchlist <- list(train=data_train, valid=data_valid)
```

I like to create the datasets and watchlist in a different code chunk to make
the code easier to iterate on in the future. I started the eta at 0.5, then
reduced it to 0.1, and then to 0.01. At 0.01 it seemed to slow down a lot but
still improve. I found that 2000 rounds seemed to be a good limit.

```{r}
model <- xgb.train(data = data_train,
                 max_depth = 4, eta = 0.01, nthread = 2,
                 nrounds = 2000, objective = "reg:linear",
                 watchlist = watchlist,
                 print_every_n = 100)
```

This is a good sign that eta has been tuned correctly and we've used
a good number of rounds: the validation RMSE has stopped improving
but has not gotten worse and the training RMSE has continued to 
improve.

```{r}
importance_matrix <- xgb.importance(model = model)
importance_matrix[,1] <- colnames(X)[as.numeric(importance_matrix[[1]]) + 1]
importance_matrix
```

So, once again trip distance is the most important. We see that hour,
dropoff location, and weekday are again important. By using non-linearities,
the location information seems even more important than we previously found.

```{r}
taxi$duration_pred_xgb <- predict(model, newdata = X)
sqrt(tapply((taxi$duration_pred_xgb - taxi$duration)^2, taxi$train_id, mean))
```

What have we learned from the boosted trees?

- Again, trip distance is the most important variable
- Hour, dropoff location, and hour are again the second most important 
- The boosted trees outperfrom the elastic net by a wide marging;
probable some non-linearities and interactions we need to handle
- Should be able to get a test RMSE of at least around 280 seconds

## What now?

When we find that the elastic net is better or at least nearly as predictive
as the gradient trees there is a lot we can do in terms of building new variables
and interactions. Linear models (remember, the elastic net is a linear model)
have none of these higher-order terms in them and need to be given them
directly in the model matrix. 

When we find that the gradient boosted trees are significantly better,
trying to custom build new variables is rarely a good idea. At best, we
will approximate the boosted trees with a linear model. A better approach
is to figure out how to either (1) improve the boosted trees model or 
(2) to construct a different model that can be blended with the boosted
trees. The first is not very interesting so let's try the second.

### Generalize additive models

Our best tool for building highly non-linear models are generalized
additive models. Here, I'll interact the longitude and latitude 
variables from both the pickup and dropoff locations, the two
most important time variables, and add a non-linear trip distance
term.

```{r}
library(gam)
model <- gam(duration ~ lo(pickup_longitude, pickup_latitude) +
                        lo(dropoff_longitude, dropoff_latitude) +
                        lo(hour, weekday) +
                        s(trip_distance),
             data = taxi,
             subset = (train_id == "train"))
```

Then, I predict the values and compare the xgboost and elastic net
models:

```{r, warning = FALSE}
taxi$duration_pred_gam <- predict(model, newdata = taxi)
sqrt(tapply((taxi$duration_pred_gam - taxi$duration)^2, taxi$train_id, mean))
```

The predictive power of this model is not as good as the boosted
trees but also does not suffer from the same degree of overfitting.

### Blending

We can usually do better by averaging together the models that we built
into a single meta-model. I tend to do this with a straightforward 
linear or logistic regression fit on only the validation set:

```{r}
model <- lm(duration ~ duration_pred_gam + duration_pred_xgb - 1,
            data = taxi,
            subset = (train_id == "valid"))
summary(model)
taxi$duration_pred <- predict(model, newdata = taxi)
```

About 85% of the final prediction comes from the boosted trees
and 13% from the gam model. 

```{r}
sqrt(tapply((taxi$duration_pred - taxi$duration)^2, taxi$train_id, mean))
```

We see that the validation error has changed only marginally, but
has improved slightly. This new meta-model is also likely more
stable and less prone to model drift and overfitting.

## Classification









