---
title: "Class 02: Sleepy Mammals and Sea Snails"
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/2017-08-31-class02/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

## Basic R

### Markdown

Now, I want to throw you all into the format for the class and
get you started with how assignments will work. This document,
as well as all of your assignments, is written in a format
called **Rmarkdown**. At the moment we are looking at the
source code; if you are familiar with markdown, it is just a
specific flavor of the basic standard.

Plain markdown is just a very simple format for marking up
texts. Unlike writing text in a word editor, there is nothing
hidden in a markdown file. If you want something to be in
italics, surround it in single stars like *this*. For bold,
use **two**. Links use the following format:
[site name](http://www.wheretogo.com). If you want to make
something look like code in the output, you can enclose
text in back ticks: `print("Hello!")`.

We can turn markdown into a number of other formats. If I
hit the *preview* button above, it will display what the
HTML version of the document looks like. I'll usually
export this HTML version and put it on the class website
as it is slightly cleaner than the raw format here.

The most powerful feature of **Rmarkdown** is that we can
intermix code into the document and actually run it in real
time. To do this, we enclose code blocks with three back ticks,
and preface the first one with r in squiggly brackets. Every
code block will get run in sequence if when we hit the Preview
or Knit buttons, but we can also run just this block by hitting
the play button to the right of the block. Here, let's add
1 and 1 together:

```{r}
1 + 1
```

The output is written directly to the document. This is the
reason I'll often use the raw format in class; it allows us
to modify the code and see the changes as they run.

Note that there are shortcuts for running a chunk of code.
To see them for your platform, look at the menu under the
Run button.

### Packages

In addition to the basic R functions that exist on start-up,
there are thousands of user-contributed packages the implement
various add-ons. To install these packages, we use the function
`install.packages`:

```{r, eval = FALSE}
install.packages("readr")
```

Once a package is installed, we also need to load it. While
installing the package only needs to be done once, we have to
load it each and every time we restart R (notice up above that
I included the option `eval = FALSE` in the code block so that
my computer does not constantly reinstall it):

```{r}
library(readr)
```

Once loaded, we can run commands from the **readr** package.
We will do this in just a moment.

### Loading Data

Let's load in four libraries that will be important for
us throughout the semester. Note that I'll turn off messages
as the packages produce quite a bit of verbose output that
we do not need to worry about. Also note that re-loading the
**readr** package has no ill-effects.

```{r, message = FALSE}
library(readr)
library(ggplot2)
library(viridis)
library(dplyr)
```

Next, let's load in a small dataset to work with today.
This data consists of the average number of hours per
day that various species are awake. We will read the data
set, like most of those in this class, directly from
my website:

```{r, message = FALSE}
msleep <- read_csv("https://statsmaths.github.io/ml_data/msleep.csv")
```

Once this code has been run, you should see the dataset
pop up in the upper right hand corner of the screen.
Clicking on it opens up a spreedsheet-like view of the
data.

Particularly important are the first three columns, as
most datasets that we work with this semester will have
these columns as well. The meaning

- the first column is just an id; don't lose it!
- the second indicates whether this is a sample where
you know the response or if have to predict the response
- the third column gives the response of interest; it
is missing whenever the second column is equal to "test"

### qplot

We will use the function `qplot` to produce simple plots
based on our data. In its most basic form, we give `qplot`
the name of the variable we want to plot and the name of
the dataset in which the variable resides.

```{r, message = FALSE}
qplot(bodywt_log, data = msleep)
```

The function is quite smart and will choose by default the
most appropriate plot to use. Here it is creating a *histogram*
because we gave it one numeric variable.

Applying `qplot` to a categorical variable gives a bar plot:

```{r}
qplot(vore, data = msleep)
```

We can also give `qplot` two variable. For instance, if we
use two continuous variables the function will construct a
scatter plot:

```{r, message = FALSE}
qplot(bodywt_log, awake, data = msleep)
```

The missing values that we are warned about are those points
where we need to produce predictions.

There is not a special name for the plot with a continuous
variable plotted versus a categorical one, but the point
should be clear (not too different from a scatter plot):

```{r}
qplot(vore, awake, data = msleep)
```

Finally, one can plot two categorical variables. It is often
not useful, but here it does help to understand the relationship
between order and diet:

```{r}
qplot(vore, order, data = msleep)
```


### Aesthetics

The `qplot` function comes from the **ggplot2** package. In
**ggplot2**, plots can be modified by changing what are known
as aesthetic mappings. Common aesthetics are: color, size, and
alpha (the opacity of the points). To change any of these to
a fixed value for all points, simply use the `I` function
in the call to `qplot`:

```{r}
qplot(vore, awake, data = msleep, color = I("gold"),
      size = I(10), alpha = I(0.3))
```

Alternatively, we can map aesthetics to variables in our dataset.
For example, let's change the color to a function of the diet
type of each mammal (note that here we do not use the `I` function):

```{r}
qplot(bodywt_log, awake, data = msleep, color = vore)
```

Color behaves differently if we use a discrete variable or a numeric
one. Notice here how a color gradient is used in place of a fixed scale:

```{r}
qplot(bodywt_log, awake, data = msleep, color = bodywt_log)
```

This last plot is particularly hard to read on the projector. It is also
difficult on my screen and for anyone with any degree of color-blindness.
The package **viridis** can help. Let's load it in (if you want to run this
line, make sure to install the package first!):

```{r}
library(viridis)
```

We now literally add a *layer* to the plot describing what color scale we
would like to make use of:

```{r}
qplot(bodywt_log, awake, data = msleep, color = bodywt_log) +
  scale_color_viridis()
```

We will make use of these aesthetics throughout the semester to explore datasets.
In particular, the color mapping allows us to look at more than two variables at
the same time.

### Best-fit line

It seems like there is a positive relationship between body
weight and hours spent awake. We can add a best-fit line to
the plot to model this relationship. Adding things to a
`qplot` graphic literally involves using the `+` sign and
adding the results of other functions.
Here we use `geom_smooth`
with an option to make the smooth strictly linear:

```{r}
qplot(bodywt_log, awake, data = msleep) +
  geom_smooth(method = "lm")
```

We can fit the exact some model analytically, rather than graphically,
by calling the `lm` function directly. Here we use an R formula: the
response variable (thing we want to predict), followed by the `~`
sign, followed by the predictor variable. As with the graphics command,
we need to indicate which dataset is being used. The output shows
the slope and intercept of the best-fit-line.

```{r}
model <- lm(awake ~ bodywt_log, data = msleep)
model
```

The function `predict` will give us the predicted values implied
by this line:

```{r}
predict(model, newdata = msleep)
```

### Accessing and adding new variables

Columns of the dataset `msleep` can be referenced by
using the dataset name followed by a dollar sign and
the name of the variable. So, for example, here are the
hours spent awake:

```{r}
msleep$awake
```

The `NA`'s correspond to the missing values we want to predict.
Similarly, we can add a new variable by simply referencing it
with the `$` and assigning it to something. Here, we add those
predictions back into the `msleep` dataset:

```{r}
msleep$awake_pred <- predict(model, msleep)
```

Click on the msleep dataset again to verify that there is a
new column of predictions. Notice that these are filled in
for all values.

### Prediction solutions

Now, we use the `select` function to pick just a few
variables from `msleep` and save this as a new dataset.
The select function takes the name of the dataset first,
followed by whatever variable names we want to keep. Here
we include just the observation id and our prediction:

```{r}
submit <- select(msleep, obs_id, awake_pred)
```

Look at the `submit` dataset and verify that it is what
you would expect.

Finally, the `write_csv` function will save a dataset
to a file. Each of the columns will be separated by a
comma (that's what the 'c' stands for in csv) and the
rows will be separated by a new line:

```{r, eval = FALSE}
write_csv(submit, "class01_submit.csv")
```

This is the kind of file I expect you to produce for
each class assignment. Notice that we've created predictions
for all of the data, even though we already know some of the
values. It is often easier to do this and avoids the mistake
of predicting on the wrong set. Of course, I'll only grade
you on the set that you were not able to see.

## GitHub Classroom

Now you should be familiar with the basics of using R to read
data, create visualizations, and produce predictions. Let's go
through and show how to put these together to produce and submit
your daily labs. I will do this interactively, but here are
the steps in case you are returning to these notes at a later
time:

1. I will send a link to your e-mail that you should follow
and accept. Assuming you have a valid GitHub account, this
will set up a repository where all of your labs for this semester
should be posted.
2. Select the lab you would like to work on (`lab01.Rmd` here)
and click on the link.
3. Now, click on the **Raw** button. This should take you to
the raw notebook file. Download this through your browser somewhere
on your computer.
4. Open this file in RStudio and work through the questions.
5. When you are finished, select the `Knit to HTML` button. This
should create an html file in the same location that you saved
the Rmd file. You should also have a file `class01_submit.csv`
in the same location.
6. Return to GitHub and drag and drop the three files into your
repository. Commit these to the repository with the **Commit changes**
button and then you are done! If dragging and dropping does not work
(only officially support on Chrome I believe), select **Upload files**
and do so manually.
7. There should already be a `lab02.Rmd` for you to work on for the
next class. New labs will show up automatically as we proceed through
the semester.

Remember that labs are due online 90 minutes before class starts
(noon on Tuesdays and Thursdays).


