\documentclass[12pt]{article}

\usepackage{fontspec}
\usepackage{geometry}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage{hyperref}

\geometry{top=1in, bottom=1in, left=1in, right=1in, marginparsep=4pt, marginparwidth=1in}

\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancyplain}
\fancyhf{}
\cfoot{\thepage\ of \pageref{LastPage}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt}

% \setromanfont [Ligatures={Common}, Numbers={OldStyle}, Variant=01,
%  BoldFont={LinLibertine_RB.otf},
%  ItalicFont={LinLibertine_RI.otf},
%  BoldItalicFont={LinLibertine_RBI.otf}
%  ]{LinLibertine_R.otf}

\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

\usepackage{xunicode}
\defaultfontfeatures{Mapping=tex-text}

\setromanfont{YaleNew}

\begin{document}

\begin{center}
{\bf MATH/CS 395: Statistical Learning, Fall 2017} \\
Tuesday, Thursday 13:30-14:45 \quad PURH G13\\
\end{center}

\bigskip

\noindent
\begin{tabular}{ l l }
{\bf Instructor:} &  {\bf Taylor Arnold} \\
E-mail: & \href{mailto:tarnold2@richmond.edu}{tarnold2@richmond.edu} \\
Office: & Jepson Hall, Rm 218 \\
Office hours: & Tuesday, Thursday 10:30-12:00 or by appointment
\end{tabular}

\vspace{0.5cm}

\textbf{Computing:} \vspace{6pt}

The focus of this course will be on applied statistics and data analysis
over symbolic mathematics. To facilitate this, nearly every class
assignment and exam will involve some form of computing.
No prior programming experience is assumed or required. \\

We will use the \textbf{R} programming environment throughout the
semester. It is freely available for all major operating systems and
is pre-installed on many campus computers. You can download it and
all supporting files for your own machine via these links:
\begin{center}
\url{https://cran.r-project.org/} \\
\url{https://www.rstudio.com/}
\end{center}
I strongly recommend using your own machine for this course, though
the lab computers in Jepson are available and contain much of the
required software as well.

\vspace{0.4cm}

\textbf{Course Website:} \vspace{6pt}

All of the materials and assignments for the course will be posted
on the class website:
\begin{quote}
\url{https://statsmaths.github.io/stat395}
\end{quote}
At the end of the semester, this version of the course
will be archived and available for your reference.

\vspace{0.4cm}

\textbf{GitHub:} \vspace{6pt}

All of your work for this semester will be submitted through GitHub,
the same platform that hosts our website. You'll need to set up a free
account, which we will cover during the week of class.

\vspace{0.4cm}

\textbf{Grades:} \vspace{6pt}

All grades in this course will be given on as a letter grade.
While occasionally possible to receive pluses / minuses or
fractional points, these will usually be given a whole letter grade. \\

I expect most students to get full marks (A) for labs and participation.
Students found to be delinquent in either will first receive a written
warning, followed by an initial 50\% reduction (C) in the respective grade,
and finally a 100\% reduction (F). \\

Your final grade will be determined by converting all grades into
a numeric scale as follows (pluses increase the number by 0.33 and minuses
decrease the number by 0.33):

\begin{center}
\begin{tabular}{c || c}
Numeric Score & Final Grade \\
\hline \hline
4 & A  \\
3 & B  \\
2 & C  \\
1 & D  \\
0 & F
\end{tabular}
\end{center}

I want to make the grading extremely transparent, so your final grade will
simply consist of taking your weighted numerical average using the following
weights:

\begin{itemize}\setlength\itemsep{0em}
\item Participation, 20\%
\item Reports, 80\%
\end{itemize}

And reading off of the following chart (grades are rounded to the
second digit):

\begin{center}
\begin{tabular}{c || c}
Numeric Score & Final Grade \\
\hline \hline
3.84 - 4.00 & A  \\
3.50 - 3.83 & A- \\
3.17 - 3.49 & B+ \\
2.84 - 3.16 & B  \\
2.50 - 2.83 & B- \\
2.17 - 2.49 & C+ \\
1.84 - 2.16 & C  \\
1.50 - 1.83 & C- \\
0.00 - 1.49 & F
\end{tabular}
\end{center}

\vspace{0.4cm}

\textbf{Reports:} \vspace{6pt}

The reports required for this class will be written documents that
mix code, graphics, and prose to provide a comprehensive analysis of
a predictive modelling task. Reports will be due at the start of class
on most Tuesdays. These must be submitted on GitHub. The format for
these reports will be described in the first week of the course.

\vspace{0.4cm}

\textbf{Exams:} \vspace{6pt}

This course has no exams, final or otherwise.

\vspace{0.4cm}

\textbf{Participation, Attendance and Late Policy:} \vspace{6pt}

You are expected to submit work on-time.
You should aim to attend all class meetings, however I am fully aware that through
the course of the semester various issues -- illness, sports, and family emergencies --
will prevent many of you from attending every class. As long as you come
prepared to most meetings of the course you will receive full credit for participation.
Likewise, attending every class but continually failing to engage in
the classroom activities will not earn full course credit.

% \newpage

% \textbf{Weekly Topics:} \vspace{6pt}

% These topics are subject to change based on the pace of the course, but
% give a good sense of roughly what we are going to cover:

% \vspace{0.5cm}

% \def\labelitemi{}
% \def\labelitemii{}

% \begin{itemize}\setlength\itemsep{0em}
% \item WEEK 01 - Introduction to R, RMarkdown, and Graphics
% \item WEEK 02 -
% \item WEEK 03 -
% \item WEEK 04 -
% \item WEEK 05 -
% \item WEEK 06 -
% \item WEEK 07 -
% \item WEEK 08 -
% \item WEEK 09 -
% \item WEEK 10 -
% \item WEEK 11 -
% \item WEEK 12 -
% \item WEEK 13 -
% \item WEEK 14 -
% \end{itemize}

\end{document}





