---
title: "Lab 04"
author: ""
output: html_notebook
---


*The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far. In this case, please restrict yourself
to multivariate linear regressions.*

# Set up

Read in the following libraries and to load the diamonds dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)

diamonds <- read_csv("https://statsmaths.github.io/ml_data/diamonds.csv")
```

The dataset gives characteristics of various types of diamonds.
Your goal is to use these to estimate the price of each diamond.

# Lab 04








# Submission

The code below assumes that you have adding a prediction named
`price_pred` to every row of the dataset.

```{r}
submit <- select(diamonds, obs_id, price_pred)
write_csv(submit, "class04_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
