---
title: "Lab 06"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

*Additionally*: For this lab, fit at least one model using the
matrix formulation we saw today with `lm.fit`. I also encourage
you to figure out a way to use the latitude and longitude
variables, though this is not a requirement.

# Set up

Read in the following libraries and to load the diamonds dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)

housing <- read_csv("https://statsmaths.github.io/ml_data/ca_pa_house_price.csv")
```

Notice that the test set here is not a random subset of the data but
has very different properties that than the other variables.

# Lab 06








# Submission

The code below assumes that you have added a prediction named
`price_pred` to every row of the dataset.

```{r}
submit <- select(housing, obs_id, price_pred)
write_csv(submit, "class06_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
