---
title: "Lab 03"
author: ""
output: html_notebook
---

## Instructions

The format for this lab is similar to the last one.
I have a section here which loads in various
libraries (if some are new, there will also be code to install
them) and then loads a dataset of interest. Your goal is to
predict the value of the third column (which will be missing
on the test set) using the techniques we have learned so far.
In this case, please restrict yourself to linear regressions,
though you may now use multivariate models.

The lab should be written as short code snippets with surrounding
text. The text should be in full sentences and should fully describe
what and why you are running the code you are running.  When there
are results, you should also give a sentence or two describing them.

Make sure to include at a minimum several plots and justify why
are picked the final model you are using.

You may not share code with other students prior to the submission
of each lab. However, discussing the labs amongst other students
is allowed. Also, copying and pasting my code from class or
prior submissions you made is **not** considered plagiarism and
is in fact highly encouraged. With permission, I may occasionally
post student-solutions to past labs. These are also fair game
for copying code from (though please do not excessively copy the
expository text). If in doubt, please ask me before submitting
results.

## Set up

Your first step should be to read in the following libraries:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
```

And to load the wine dataset:

```{r}
wine <- read_csv("https://statsmaths.github.io/ml_data/wine.csv")
```

The dataset gives characteristics of various types of wine.
Your goal is to use these to estimate the quality of the wine.

Your code and text should go in the section below (feel free to create
sub-sections with the `###` command at the start of line, though please
note that this is not required):

## Lab 03

THIS IS WHERE YOUR CODE AND TEXT SHOULD GO (and please remove this line!)




## Submission

The code below assumes that you have adding a prediction named
`quality_pred` to every row of the `wine` dataset.

```{r}
submit <- select(wine, obs_id, quality_pred)
write_csv(submit, "class03_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to GitHub.
