---
title: "Lab 18"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

One thing that you are not allowed to use are neural networks.
We will see how to use this with images next class.

# Set-up

Read in the following libraries and to load the metadata about
the flowers:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(stringi)

f17 <- read_csv("https://statsmaths.github.io/ml_data/flowers_17.csv")
```

To get the actual data for the images, you'll have to download the following
file.

- https://drive.google.com/open?id=0B6_9OUDRaPQsc21sZXQ2Q250bFk

Once these are downloaded, you'll have to run something like this:

```{r}
x64 <- read_rds("~/Desktop/flowers_17_x64.rds")
```

If, for example, the file sits on your Desktop and you have a Mac. If you have
trouble with any of this, please let me know as soon as possible.


# Lab 18











# Submission

The code below assumes that you have added a prediction named
`class_pred` to every row of the dataset.

```{r}
submit <- select(flowers, obs_id, class_pred)
write_csv(submit, "class18_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
