---
title: "Lab 11"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

For this lab, please only use neural networks.

# Set up

Read in the following libraries and to load the crimes dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)

crimes <- read_csv("https://statsmaths.github.io/ml_data/chi_crimes_12.csv")
```

The names of the crimes are, in order:

```{r}
crimes_list <- c("theft", "battery", "criminal damage", "narcotics",
"other offense", "assault", "burglary", "motor vehicle theft", "robbery",
"deceptive practice", "criminal trespass", "prostitution")
```

Make sure your predictions are all an integer from 1 to 12.

```{r}
if (!require("keras")) {
  install.packages("keras")
}
```

# Lab 11






# Submission

The code below assumes that you have added a prediction named
`crime_type_pred` to every row of the dataset.

```{r}
submit <- select(crimes, obs_id, crime_type_pred)
write_csv(submit, "class11_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
