---
title: "Lab 16"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

One thing that you are not allowed to use are character shingles.
I will give you the parsed text of the blog posts in case you
would like to use part of speech codes or other filtering
schemes to construct the data matrix

# Set-up

Read in the following libraries and to load the amazon products:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(stringi)
library(tokenizers)
library(smodels)

blog <- read_csv("https://statsmaths.github.io/ml_data/blog_age.csv")
```

You can also load the parsed data from:

```{r}
blog <- read_csv("https://statsmaths.github.io/ml_data/blog_age_anno.csv.gz")
```


# Lab 16











# Submission

The code below assumes that you have added a prediction named
`teen_pred` to every row of the dataset.

```{r}
submit <- select(blog, obs_id, teen_pred)
write_csv(submit, "class16_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
