---
title: "Lab 12"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

**One special rule this time: You are absolutely forbidden to
create new variables by hand. If I find, for example, that you
created a "flush" column or a column "max card" you'll get a zero
for the lab.**

# Set up

Read in the following libraries and to load the crimes dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)

poker <- read_csv("https://statsmaths.github.io/ml_data/poker_hands.csv")
```

Your assignment will be scored using RMSE.

# Lab 12











# Submission

The code below assumes that you have added a prediction named
`quality_pred` to every row of the dataset.

```{r}
submit <- select(poker, obs_id, quality_pred)
write_csv(submit, "class12_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
