---
title: "Lab 13"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

# Set up

You may need the tokenizers and smodels libraries if you don't
already have them:

```{r}
if(!require("tokenizers")) {
  install.packages("tokenizers")
}
if(!require("smodels")) {
  if(!require("devtools")) {
    install.packages("devtools")
  }
  devtools::install_github("statsmaths/smodels")
}
```

Read in the following libraries and to load the amazon products:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(stringi)
library(tokenizers)
library(smodels)

president <- read_csv("https://statsmaths.github.io/ml_data/presidents_5.csv")
```

Your assignment will be scored using accuracy. Make sure to submit class
labels of either 1, 2, 3, 4 or 5.

# Lab 14











# Submission

The code below assumes that you have added a prediction named
`class_pred` to every row of the dataset.

```{r}
submit <- select(president, obs_id, class_pred)
write_csv(submit, "class14_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
